---
title: "Jackknife"
author: "Ozan"
output: html_document
---

# Project Goals

The goal of this project is to work through building functions around the Jackkinfe. We will start with some simple functions and then determine a pattern to build a more general function in the end. 

## Jackknife of the Mean!

- Have an estimator $\hat{\theta}$ of parameter $\theta$  
  want the standard error of our estimate, $se_{\hat{\theta}}$
- The jackknife approximation:
    + omit case $i$, get estimate $\hat{\theta}_{(-i)}$
    + Take the variance of all the $\hat{\theta}_{(-i)}$
    + multiply that variance by $\frac{(n-1)^2}{n}$ to get $\approx$ variance of $\hat{\theta}$
- then $se_{\hat{\theta}}=$ square root of that variance

**PHP 2560 Only** (Why $(n-1)^2 / n$?  Think about just getting the standard
    error of the mean)

  We use $(n-1)^2 / n$ to scale the variance estimations, which comes from SE of the mean formula.

Write a function called `mean.jackknife` that takes argument `a_vector` and returns a jackknife estimate of the standard error.

```{r, eval=FALSE}
mean.jackknife <- function(a_vector) {
  n <- length(a_vector)
  jk.ests <- vector(length = n)
  for (removed in seq_along(jk.ests)) {
    jk.ests[removed] <- mean(a_vector[-removed])
  }
  var.ests <- var(jk.ests)
  jk.var <- ( (n-1)^2 / n) * var.ests
  jk.stderr <- sqrt(jk.var)
  return(jk.stderr)
}
```

#### Jackknife for the mean Example: test your code here

```{r, eval = FALSE}
some_normals <- rnorm(100, mean=7 ,sd=5)
mean(some_normals)
(formula_se_of_mean <- sd(some_normals) / sqrt(length(some_normals)))
all.equal(formula_se_of_mean, mean.jackknife(some_normals))
```

##Jackknife for Gamma Parameters

The following function is a way to calculate the method of moments estimators for the gamma distribution:

```{r, eval=FALSE}
gamma.est <- function(the_data) {
  m <- mean(the_data)
  v <- var(the_data)
  a <- m^2 / v
  s <- v / m
  return(c(a = a,s = s))
}
```

###Jackknife for Gamma Parameters Function

Write a function called `gamma.jackknife` that takes argument `a_vector` and returns jackknife standard error estimates on the gamma parameters. 

```{r, eval=FALSE }
gamma.jackknife <- function(a_vector) {
  n <- length(a_vector)
  jk.ests <- matrix(NA, nrow = 2, ncol = n)
  rownames(jk.ests) = c("a", "s")
  for (removed in 1:n) {
    gamma.ests <- gamma.est(a_vector[-removed])
    jk.ests["a", removed] <- gamma.ests["a"]
    jk.ests["s", removed] <- gamma.ests["s"]
  }
  var.ests <- apply(jk.ests, 1, var)
  jk.vars <- ( (n-1)^2 / n) * var.ests
  jk.stderrs <- sqrt(jk.vars)
  return(jk.stderrs)
}
```

###Jackknife for Gamma Parameters Example

```{r, eval=FALSE}
input <- rgamma(1000, shape=0.1, scale = 10)
gamma.est(input)
gamma.jackknife(input)
```

##Jackknife for linear regression coefficients

Write a function called `jackknife.lm` that takes arguments `df`, `formula` and `p` and returns jackknife standard error estimates on the coefficients of a linear regression model. 

```{r, eval=FALSE}
jackknife.lm <- function(df, formula, p) {
  n <- nrow(df)
  jk.ests <- matrix(NA, nrow = p, ncol = n)
  for (removed in 1:n) {
    coefs <- lm(formula, data = df[-removed, ])$coefficients
    jk.ests[ , removed] <- coefs
  }
  var.ests <- apply(jk.ests, 1, var)
  jk.var <- ( (n-1)^2 / n) * var.ests
  jk.stderr <- sqrt(jk.var)
  return(jk.stderr)
}
```

###Jackknife for linear regression coefficients Example

```{r, eval=FALSE}
output <- 1.2 + 0.6*input +  rnorm(1000, 0, 2.1)
data <- data.frame(output,input)
my.lm <- lm(output~input, data=data)
coefficients(my.lm)
# "Official" standard errors
sqrt(diag(vcov(my.lm)))
jackknife.lm(df=data,formula="output~input",p=2)
```

##Refactoring the Jackknife

- Omitting one point or row is a common sub-task

- The general pattern:

```
figure out the size of the data
for each case
   omit that case
   repeat some estimation and get a vector of numbers
take variances across cases
scale up variances
take the square roots
```
- Refactor by extracting the common "omit one" operation

- Refactor by defining a general "jackknife" operation

###The Common Operation

- _Problem_: Omit one particular data point from a larger structure

- _Difficulty_: Do we need a comma in the index or not?

- _Solution_: Works for vectors, lists, 1D and 2D arrays, matrices, data frames:

## Goal: 

- Make the function select the correct dimensions
    - length for a 1d object
    - number of rows for 2d
- Write a function `omit.case` that omits a point given the data and returns the data minus that point. Make sure it can handle higher dimensions. 

```{r, eval=FALSE}
omit.case <- function(holder, point) {
  dims <- dim(holder)
  if (is.null(dims) | (length(dims) == 1)) {
    return(holder[-point])
  }
  return(holder[-point, ])
}
```

- Write a function `omit_and_est` that takes the data with an omitted point and returns whatever function your estimator does.  
```{r,eval=FALSE}
omit_and_est <- function(omit) {
  estimator(omit.case(holder, omit))
}
```

```{r, eval=FALSE}
jackknife <- function(estimator, holder) {
  
  # Define n to be lenth or number of rows here
  n <- ifelse(is.null(dim(holder)), length(holder), nrow(holder))
  omit_and_est <- function(omit) {
    estimator(omit.case(holder, omit))
  }
  jk.ests <- matrix(sapply(1:n, omit_and_est), ncol = n)
  var.reests <- apply(jk.ests, 1, var)
  jk.var <- ( (n-1)^2 / n) * var.reests
  jk.stderr <- sqrt(jk.var)
  return(jk.stderr)
}
```

---

###It works

```{r, eval=FALSE}
jackknife(estimator = mean, holder = some_normals)
all.equal(jackknife(estimator = mean, holder = some_normals),
          mean.jackknife(some_normals))
```

---

```{r, eval=FALSE}
all.equal(jackknife(estimator=gamma.est,holder = data$input),
          gamma.jackknife(data$input))
```

---

```{r, eval=FALSE}
all.equal(jackknife(estimator=gamma.est, holder = data$input),
          gamma.jackknife(data$input), check.names=FALSE)
```

```{r, eval=FALSE}
est.coefs <- function(the_data) {
  return(lm(output~input, data = the_data)$coefficients)
}
est.coefs(data)
all.equal(est.coefs(data), coefficients(my.lm))
```

---

```{r, eval=FALSE}
jackknife(estimator=est.coefs, holder = data)
all.equal(jackknife(estimator=est.coefs,holder = data),
          jackknife.lm(df = data, formula="output~input",p=2))
```

##Further Refactoring of jackknife()

The code for `jackknife()` is still a bit clunky:
- Ugly `if-else` for finding `n`
- Bit at the end for scaling variances down to standard errors

- write a function that calculates the `n` needed for the above code:

```{r, eval=FALSE}
data_size <- function(holder) {
  n <- ifelse(is.null(dim(holder)), length(holder), nrow(holder))
}
```

- Write a function that calculate the variance of all the estimates and returns the standard error

```{r, eval=FALSE}
scale_and_sqrt_vars <- function(jk.ests, n) {
  var.reests <- apply(jk.ests, 1, var)
  jk.var <- ( (n-1)^2 / n) * var.reests
  jk.stderr <- sqrt(jk.var)
  return(jk.stderr)
}
```

###Now invoke those functions

```{r, eval=FALSE}
jackknife <- function(estimator, holder) {
  n <- data_size(holder)
  omit_and_est <- function(omit) {
    estimator(omit.case(the_data,omit))
  }
  jk.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  return(scale_and_sqrt_vars(jk.ests, n))
}
```

